### 2024-04-09 paper/paper.qmd
modelsummary ran in paper.qmd

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

inaug_model <-
  readRDS(file = here::here("models/inaug_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "Inauguration model" = inaug_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-04-09 paper/paper.qmd
tested alternative model designs, abandoned new approaches and sticking to the initial version. Updated paper title, added TOC and created a plot illustrating data size difference between treatment and control

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "Effect of Election on Currency Fluctuation"
+subtitle: "Analysis of USD Exchange Rate Fluctuation during Inaugurtion week"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Sandy Yu
+thanks: "Code and data are available at: https://github.com/Jingying-yu/election-season-and-currency-fluctuation.git"
 date: today
@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(tidyverse)
-library(palmerpenguins)
+library(readr)
+library(rstanarm)
+library(knitr)
+library(modelsummary)
+library(ggplot2)
 ```
@@
 
+```{r}
+exchange_inaug |>
+  ggplot(aes(
+    x = date,
+    y = exchange_rate
+  )) +
+  geom_point(alpha = 0.005) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 0),
+    method = "lm",
+    color = "grey",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1974-08-06" & date <= "1974-08-12"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1977-01-17" & date <= "1977-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1981-01-17" & date <= "1981-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1981-01-17" & date <= "1981-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1989-01-17" & date <= "1989-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1993-01-17" & date <= "1993-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2001-01-17" & date <= "2001-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2009-01-17" & date <= "2009-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2017-01-17" & date <= "2017-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2021-01-17" & date <= "2021-01-23"),
+    method = "lm",
+    color = "blue",
+    formula = "y ~ x"
+  )
+
+```
+
 # Discussion

### 2024-04-09 scripts/02-data_cleaning.R

@@ -66,9 +66,17 @@ for(i in 1:nrow(cleaned_inauguration)) {
   }
 }
 
-# Separate the exchange rate into 2 files, one containing the rate not contain inauguration week 
-# Where the date of the inauguration is the 4th day, that is inauguration date +-3 days is the inauguration week
-# The other file containing ONLY exchange rates during inauguration week.
+#alternative approach, ABANDONED
+#inaug_period_exchange <- exchange_inaug |> filter(inauguration_period == 1)
+#for(i in 1:nrow(cleaned_inauguration)) {
+  # Define the date range for each inauguration date
+#  date_range <- seq(cleaned_inauguration$inauguration_date[i] - 3,
+#                    cleaned_inauguration$inauguration_date[i] - 1,
+#                    by = 'day')
+#  date_indices <- which(inaug_period_exchange$date %in% date_range)
+#  inaug_period_exchange$inauguration_period[date_indices] <- 0  
+#}
+
 
 
 
@@ -81,11 +89,5 @@ write_parquet(cleaned_exchange_rate, "data/analysis_data/cleaned_exchange_rate.p
 # write_csv(cleaned_inauguration, "data/analysis_data/cleaned_inauguration.csv")
 write_parquet(cleaned_inauguration, "data/analysis_data/cleaned_inauguration.parquet")
 
+write_csv(exchange_inaug, "data/analysis_data/exchange_inaug.csv")
 write_parquet(exchange_inaug, "data/analysis_data/exchange_inaug.parquet")
-
-
-#write_csv(inaug_exchange_rate, "data/analysis_data/inaug_exchange_rate.csv")
-#write_parquet(inaug_exchange_rate, "data/analysis_data/inaug_exchange_rate.parquet")
-
-#write_csv(exchange_rate_remaining, "data/analysis_data/exchange_rate_remaining.csv")
-#write_parquet(exchange_rate_remaining, "data/analysis_data/exchange_rate_remaining.parquet")

### 2024-04-09 scripts/04-model.R

@@ -11,10 +11,12 @@
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(readr)
 library(arrow)
 
 #### Read data ####
-exchange_inaug <- read_parquet("data/analysis_data/exchange_inaug.parquet")
+#exchange_inaug <- read_parquet("data/analysis_data/exchange_inaug.parquet")
+inaug_period_exchange <- read_csv("data/analysis_data/inaug_period_exchange.csv", show_col_types = FALSE)
 
 ### Model data ####
 inaug_model <-
@@ -25,9 +27,21 @@ inaug_model <-
     prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
     prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 21
+   seed = 21
   )
 
+
+# Attempt alternative RDD approach, ABANDONED
+#inaug_period_model <-
+#  stan_glm(
+#    formula = exchange_rate ~ date + inauguration_period,
+#    data = inaug_period_exchange,
+#    family = gaussian(),
+#    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+#    prior_aux = exponential(rate = 1, autoscale = TRUE),
+#   seed = 21
+#  )
 
 #### Save model ####
 saveRDS(
@@ -35,4 +49,10 @@ saveRDS(
   file = "models/inaug_model.rds"
 )
 
+
+# alternative approach, ABANDONED
+#saveRDS(
+#  inaug_period_model,
+#  file = "models/inaug_period_model.rds"
+#)
 

### 2024-04-09 paper/paper.qmd
added legend to explain plot

--- 2024-04-09 paper/paper.qmd
@@
   statistic = "mad",
-  fmt = 2
+  fmt = 2,
+  metrics = c("R2")
 )
+```
+
+```{r}
+colors <- c("No Inauguration Announcement" = "yellow", "Inauguration Announcement" = "blue")
+
+exchange_inaug |>
+  ggplot(aes(
+    x = date,
+    y = exchange_rate
+  )) +
+  geom_point(alpha = 0.005) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 0),
+    method = "lm",
+    aes(color = "No Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1974-08-06" & date <= "1974-08-12"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1977-01-17" & date <= "1977-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1981-01-17" & date <= "1981-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1981-01-17" & date <= "1981-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1989-01-17" & date <= "1989-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "1993-01-17" & date <= "1993-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2001-01-17" & date <= "2001-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2009-01-17" & date <= "2009-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2017-01-17" & date <= "2017-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) +
+  geom_smooth(
+    data = exchange_inaug |> filter(inauguration_period == 1 & date >= "2021-01-17" & date <= "2021-01-23"),
+    method = "lm",
+    aes(color = "Inauguration Announcement"),
+    formula = "y ~ x"
+  ) + scale_color_manual(values = colors) +
+  labs(
+    x = "Date",
+    y = "USD v.s. CAD Exchange Rate",
+    title = "Exchange Rate from 1974 to 2024",
+    color = "Period"
+  ) + theme(legend.position = "bottom")
+
 ```

### 2024-04-10 scripts/00-simulate_data.R
finished simulating data and added preamble to model script

@@ -9,10 +9,37 @@
 
 #### Workspace setup ####
 library(tidyverse)
-
+library(lubridate)
 
 #### Simulate data ####
 
+# Set seed for reproducibility
+set.seed(21)
+
+# Generate the sequence of dates from 1970-01-01 to 2024-01-01
+dates <- seq(from = as.Date("1970-01-01"), to = as.Date("2024-01-01"), by = "day")
+
+# Initialize the inauguration_period column with zeros
+inauguration_period <- rep(0, length(dates))
+
+# Set inauguration_period to 1 for each inauguration week every four years
+for(year in seq(1970, 2024, by = 4)) {
+  inauguration_start <- as.Date(paste(year, "-01-17", sep = ""))
+  inauguration_end <- inauguration_start + 6
+  inauguration_period[dates >= inauguration_start & dates <= inauguration_end] <- 1
+}
+
+# Generate random exchange_rate values with normal distribution
+# Make sure the values are clipped to the range [0.80, 1.55]
+exchange_rate <- rnorm(length(dates), mean = 1, sd = 0.1)
+exchange_rate <- pmin(1.55, pmax(0.80, exchange_rate))
+
+# Create the data frame
+simulated_data <- data.frame(
+  date = dates,
+  exchange_rate = exchange_rate,
+  inauguration_period = inauguration_period
+)
 
 
 

### 2024-04-10 scripts/04-model.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: create bayesian model based on cleaned data
+# Author: Sandy Yu
+# Date: 2 April 2024
+# Contact: jingying.yu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: downloaded and cleaned data, understand what model to use
+
 
 
 #### Workspace setup ####

### 2024-04-10 paper/paper.qmd
tested model in the appendix

@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
 ```
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+$$y_i|\mu_i, \sigma \sim \mbox{Normal}(\mu_i, \sigma)$$
+
+$$\mu_i = \alpha + \beta_i + \omega_i$$
+
+$$\alpha \sim \mbox{Normal}(0, 2.5)$$
+
+$$\beta \sim \mbox{Normal}(0, 2.5)$$
+
+$$\omega \sim \mbox{Normal}(0, 2.5)$$
+
+$$\sigma \sim \mbox{Exponential}(1)$$
+
+Where:
+
+-   $y_i$ is the daily exchange rate of USD v.s. CAD (base unit is CAD = 1)
+
+-   $\beta_i$ is a dummy variable (value is either 0 or 1) indicating whether the date is within the inauguration week (see data section for definition of inauguration week).
+If date is within the inauguration week, then the variable has a value of 1, and 0 otherwise.
+
+-   $\omega_i$ is a dummy variable indicating whether a change of political party occurred in the corresponding election season.
+Each inauguration week receives identical value (the whole week is either all 1 or all 0).
+Value of 1 indicates that the elected president's political party is different from last season's president's, 0 otherwise.
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
+## Explaination for small R^2
 ```{r}
@@
 
-# Discussion
+# Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+## Expectation of the population
 
@@
 
-## Second discussion point
+## Change of Political Party
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
+We can see that the regression outcome from our data and the simulation created by the posterior distribution are not closely matched.
+This is because our predictor variables only contribute to a small section of the data (40 out of 13358 observations has inauguration_period == 1, 30 out of 13358 has change_party == 1).
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
+We can see that the estimation parameters shifts minimally after taking data into account.
+This suggests that good prior parameters were set.
 
@@
 
-pp_check(first_model) +
+pp_check(inaug_model) +
   theme_classic() +
@@
 
-posterior_vs_prior(first_model) +
+posterior_vs_prior(inaug_model) +
   theme_minimal() +
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+@fig-stanareyouokay-1 is a trace plot.
+Trace plots are often used to diagnose abnormalities of models.
+For both plots, lines are oscillating vertically and maintains an overall horizontal trend.
+This suggests that there is nothing out of the ordinary for our model.
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-stanareyouokay-2 is a Rhat plot.
+Rhat plot is a valuable tool for assessing convergence in Bayesian regression models, helping ensure the reliability and validity of the model's inference results.
+Both both plots, all data points are close to 1.
+This suggest that there is nothing out of the ordinary for our model.
 
@@
 
-plot(first_model, "trace")
+plot(inaug_model, "trace")
 
-plot(first_model, "rhat")
+plot(inaug_model, "rhat")
 ```

### 2024-04-12 paper/paper.qmd
created a datasheet text file and pasted in a template, will change content in future. Also revised structure of the paper

@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+## Data on Exchange Rate
 
@@
 
-Talk more about it.
+## Data on Election Results
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+### Selection of Data Source
 
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
 ```
 
-Talk way more about it.
+### Construction of Inauguration Week
 

### 2024-04-14 paper/paper.qmd
added file location for datasheet in the READMe and answered datasheet questions up to "Collection Process"

@@
 library(ggplot2)
+library(arrow)
 ```
@@
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-exchange
+#| fig-cap: Exchange Rate from 1971 to 2024
 #| echo: false
 
+exchange_inaug <- read_csv(here::here("data/analysis_data/exchange_inaug.csv"), show_col_types = FALSE)
+
+exchange_inaug |>
+  ggplot(aes(
+    x = date,
+    y = exchange_rate
+  )) +
+  geom_line() +
+  labs(
+    x = "Year",
+    y = "Exchange Rate: USD to CAD (Base Currency: USD)",
+    title = "Exchange Rate from 1971 to 2024"
+  )
 ```
@@
 
--   $y_i$ is the daily exchange rate of USD v.s. CAD (base unit is CAD = 1)
+-   $y_i$ is the daily exchange rate of USD v.s. CAD (base unit is USD = 1)
 

### 2024-04-14 paper/datasheet.qmd
fixed rendering issues and changed directory for datasheet

---
title: "Datasheet for 'exchange_inaug'"
subtitle: "A dataset containing exchange rate information from 1971 to 2024, concerning US Presidential Election"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/election-season-and-currency-fluctuation.git"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
---

Extract of the questions from @gebru2021datasheets.

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of the relationship between US presidential election result announcement and USD exchange rate fluctuation. We were unable to find a publicly available dataset in a structured format that had all the necessary elements that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Sandy Yu, while studying at the University of Toronto as an undergraduate student.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - No funding was received for this project.
4. *Any other comments?*
    - No

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each row of the final dataset (exchange_inaug) is an date with the range of 1971 - 2024, sorted ascending. Each date has its corresponding daily exchange rate, a binary variable indicating whether the date is within the week of a US Presidential result announcement day (the dataset considers the date of the official inauguration as the announcement day for the new president), and a binary variable indicating whether there has been a change of political party this election season.
2. *How many instances are there in total (of each type, if appropriate)?*
	- 13,358 dates have been recorded.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset takes the common date between maximum date range for FRED's exchange rate and past US inauguration dates.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of the respective date, exchange rate in numeric value, and two binary indicators detailed above.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- Yes, the date values are unique for each instance.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- Exchange rates are not available for all dates within the range, some dates have been droped due to exchange rate being the response variable (estimate).
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- Yes, through the unique dates. Each date within the range will appear no more than once in the dataset.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No
9. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- Exchange Rate between CAD & USD gathered from [FRED] (https://fred.stlouisfed.org/series/DEXCAUS)
	- US Presidential Election Result data gathered from [United States Census Bureau](https://www.census.gov/content/dam/Census/library/visualizations/2021/comm/inauguration-day.pdf), cross checked with data from [Frank LaRose Ohio Secretary of State](https://www.ohiosos.gov/elections/election-results-and-data/historical-election-comparisons/presidents-of-the-united-states-of-america/) website to ensure accuracy.
10. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No, all data were gathered from public sources.
11. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No
12. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- No
13. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- The dataset does not concern any individual.
14. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No
15. *Any other comments?*
	- No

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2024-04-14 paper/paper.qmd
fixed rendering issues and changed directory for datasheet

@@
 ### Construction of Inauguration Week
+$$Day1 \hspace{5mm} Day2 \hspace{5mm}  Day3 \hspace{7mm}  Inauguration\ Day \hspace{7mm} Day5 \hspace{5mm} Day6 \hspace{5mm} Day7$$
 
@@
 ```{r}
+#| label: fig-size_visual
+#| fig-cap: Estimate average change in Exchange Rate
+#| echo: false
+#| warning: false
+#| message: false
 colors <- c("No Inauguration Announcement" = "yellow", "Inauguration Announcement" = "blue")

### 2024-04-14 paper/paper.qmd
added two tables in the data section for better visualization

@@
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: tbl-inauguration
+#| tbl-cap: Sample of past inauguration dates information
 #| echo: false
-#| warning: false
-#| message: false
+cleaned_inauguration <- read_csv(here::here("data/analysis_data/cleaned_inauguration.csv"), show_col_types = FALSE)
 
+knitr::kable(cleaned_inauguration, col.names = c("President", "Party", "Inauguration Date", "Change Party"), align = c("c", "c", "c", "c"))
 ```
@@
 ### Construction of Inauguration Week
-$$Day1 \hspace{5mm} Day2 \hspace{5mm}  Day3 \hspace{7mm}  Inauguration\ Day \hspace{7mm} Day5 \hspace{5mm} Day6 \hspace{5mm} Day7$$
+$$Day1 \hspace{5mm} Day2 \hspace{5mm}  Day3 \hspace{7mm}  \textbf{Inauguration Day} \hspace{7mm} Day5 \hspace{5mm} Day6 \hspace{5mm} Day7$$
+```{r}
+#| label: tbl-exchange_inaug
+#| tbl-cap: Sample of the final dataset used for analysis
+#| echo: false
+sample_tbl <- exchange_inaug |> filter(date >= "1977-01-12" & date <= "1977-01-18")
+
+knitr::kable(sample_tbl, col.names = c("Date", "Exchange Rate", "Inauguration Period", "Change Party"), align = c("c", "c", "c", "c"))
+```
 

### 2024-04-15 paper/datasheet.qmd
completed dataset

@@
 date-format: "D MMMM YYYY"
+format: pdf
+bibliography: references.bib
 ---
@@
 Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
-	- Exchange Rate between CAD & USD gathered from [FRED] (https://fred.stlouisfed.org/series/DEXCAUS)
-	- US Presidential Election Result data gathered from [United States Census Bureau](https://www.census.gov/content/dam/Census/library/visualizations/2021/comm/inauguration-day.pdf), cross checked with data from [Frank LaRose Ohio Secretary of State](https://www.ohiosos.gov/elections/election-results-and-data/historical-election-comparisons/presidents-of-the-united-states-of-america/) website to ensure accuracy.
+
+-   Exchange Rate between CAD & USD gathered from @FRED
+
+-   US Presidential Election Result data gathered from @Census, cross checked with data from @Ohio.
+
 10.
@@
 If so, please describe how.*
-	- TBD
+    -   Data on exchange rate is pulled from @FRED website linked above
+    -   Inauguration date and other political data is hand-gathered from multiple government websites (for cross-reference) linked above. (@Census & @Ohio)
 2.
@@
 How were these mechanisms or procedures validated?*
-	- TBD
+    -   FRED data (exchange rate) is downloaded through download link directly into RStudio using coding language R.
+    -   Inauguration date and other political data are hand gathered into an excel sheet due to the format of the pdf and website.
 3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
-	- TBD
+    -   The downloaded exchange rate data has the maximum number of observations available on the @FRED website
+    -   The inauguration data is not a sample
 4.  *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
-	- TBD
+    -   Sandy Yu, for the purpose of personal interest, therefore not financially compensated
 5.
@@
 If not, please describe the timeframe in which the data associated with the instances was created.*
-	- TBD
+    -   Data gathered in March 2024
 6.
@@
 If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+    -   No
 7.  *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
-	- TBD
+    -   Third party websites
 8.
@@
 If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
-	- TBD
+    -   No, but all the data are open-sourced
 9.
@@
 If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
-	- TBD
+    -   Open source data
 10.
@@
 If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
-	- TBD
+    -   Not applicable
 11.
@@
 If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
-	- TBD
+    -   No
 12. *Any other comments?*
-	- TBD
+    -   No
 
@@
 If not, you may skip the remaining questions in this section.*
-	- TBD
+    -   Yes cleaning of the data was done.
 2.
@@
 If so, please provide a link or other access point to the "raw" data.*
-	- TBD
+    -   Yes
 3.  *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
-	- TBD
+    -   R and Excel were used
 4.  *Any other comments?*
-	- TBD
+    -   No
 
@@
 1.  *Has the dataset been used for any tasks already? If so, please provide a description.*
-	- TBD
+    -   No, the dataset is specifically put together for this paper.
 2.  *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
-	- TBD
+    -   No
 3.  *What (other) tasks could the dataset be used for?*
-	- TBD
+    -   Potentially, but none I can think of right now
 4.
@@
 Is there anything a dataset consumer could do to mitigate these risks or harms?*
-	- TBD
+    -   No
 5.  *Are there tasks for which the dataset should not be used? If so, please provide a description.*
-	- TBD
+    -   No
 6.  *Any other comments?*
-	- TBD
+    -   No
 
@@
 If so, please provide a description.*
-	- TBD
+    -   No
 2.  *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
-	- TBD
+    -   Data will be available on Github
 3.  *When will the dataset be distributed?*
-	- TBD
+    -   data is available now
 4.
@@
 If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
-	- TBD
+    -   No. MIT license.
 5.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
-	- TBD
+    -   None that are known
 6.
@@
 If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
-	- TBD
+    -   None that are known
 7.  *Any other comments?*
-	- TBD
+    -   No
 
@@
 1.  *Who will be supporting/hosting/maintaining the dataset?*
-	- TBD
+    -   Sandy Yu
 2.  *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
-	- TBD
+    -   jingying.yu\@mail.utoronto.ca
 3.  *Is there an erratum? If so, please provide a link or other access point.*
-	- TBD
+    -   No, the dataset is just updated.
 4.
@@
 If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
-	- TBD
+    -   Unlikely
 5.
@@
 If so, please describe these limits and explain how they will be enforced.*
-	- TBD
+    -   Not applicable
 6.  *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
-	- TBD
+    -   The updated dataset is the first version and likely the last
 7.
@@
 If so, please provide a description.*
-	- TBD
+    -   Pull request on GitHub.
 8.  *Any other comments?*
-	- TBD
+    -   No
 

### 2024-04-15 scripts/02-data_cleaning.R

@@ -83,10 +83,10 @@ for(i in 1:nrow(cleaned_inauguration)) {
 
 
 #### Save data ####
-# write_csv(cleaned_exchange_rate, "data/analysis_data/cleaned_exchange_rate.csv")
+write_csv(cleaned_exchange_rate, "data/analysis_data/cleaned_exchange_rate.csv")
 write_parquet(cleaned_exchange_rate, "data/analysis_data/cleaned_exchange_rate.parquet")
 
-# write_csv(cleaned_inauguration, "data/analysis_data/cleaned_inauguration.csv")
+write_csv(cleaned_inauguration, "data/analysis_data/cleaned_inauguration.csv")
 write_parquet(cleaned_inauguration, "data/analysis_data/cleaned_inauguration.parquet")
 
 write_csv(exchange_inaug, "data/analysis_data/exchange_inaug.csv")

### 2024-04-15 paper/paper.qmd
completed abstract and added to intro

@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper analyzes the effect of US Presidential Election Announcement and a change of political party on the USD exchange rate.
+Daily exchange rates gathered from FRED and past presidential inauguration dates from multiple government websites are used in a generalized linear regression model to estimate the effects.
+We find that election result announcement has an average negative effect on USD exchange rate whereas a change of political party contributes positively with a slightly larger magnitude.
+Our result suggests that exchange USD for another currency closly after a Presidential election with a change of political party could be profitable."
 format: pdf
@@
 The remainder of this paper is structured as follows. @sec-data....
+
+Analyses and findings in this paper are structured into several sections: @sec-data -- Data, @sec-model – Model, @sec-results -- Results, and @sec-discussion -- Discussion.
+The Data section examines all datasets and variables kept for analysis, followed by an explanation of their data cleaning processes.
+The Model section defines linear models used for further analysis, explain its components, and presents model justifications.
+The Result section focuses on visualizing and presenting the model results through data presented in Data section.
+The Discussion section further evaluate the interpretations behind the model results presented in the previous section, and touches on any weaknesses and next steps.
 

### 2024-04-15 paper/paper.qmd
updated reference.bib and finshed introduction & model section of paper

@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+USD is one of the most popular currency in circulation today.
+As a foreign buyer of USD product, the final price can vary greatly depending on the exchange rate between one's local currency and USD.
+Fluctuation of exchange rate can be attributed to multiple different factors.
+For example the difference in interest rate between two countries, inflation rate, etc.
+This paper will focus on one particular aspect that contributes to the change in USD exchange rate: the expectation of the populous on the future of the United States, perceived through the results of the US Presidential Election.
 
-The remainder of this paper is structured as follows. @sec-data....
+This paper seeks to explore the effect that the US Presidential Election result announcements have on USD interest rate.
+USD v.s.
+CAD exchange rate from 1971 to 2024 is pulled from the @FRED website, and past presidential election results and inauguration dates are gathered and double-checked through @Cencus and @Ohio.
+The estimand of the paper is how the announcement and inauguration of a new US president, and how a change in political party, influence the USD exchange rate.
+
+The result of our analysis indicates that the announcement and inauguration of a new US president has an average negative effects on USD exchange rate within the time frame of 1971 to 2024.
+If a change of political party has occured for the new election season, where the leading party this season is different from the previous, then an average positive effect occurs.
+The magnitude of effect for the two factors suggest that, if the new president belongs to a different party than the previous president, then the inauguration anncouncement will have a net positive effect on USD exchange rate.
 
@@
 # Data {#sec-data}
+
+The USD exchange rate data, denoted DEXCAUS [@Federal_Reserve], is obtained through the FRED website [@FRED].
+The source of the dataset is the US Board of Governors of the Federal Reserve System [@Federal_Reserve].
+The past US president inauguration dates are pulled from the US Census Bureau [@Census] and cross-referenced with Frank LaRose Ohio Secretary of State [@Ohio].
+
+Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], rstanarm [@rstanarm], arrow [@arrow], ggplot2 [@ggplot2], and knitr [@knitr], with additional help from Microsoft Excel [@citeExcel].
+Detailed description of each dataset can be found in the subsections below.
 
@@
   )) +
-  geom_line() +
+  geom_line() + theme_minimal() +
   labs(
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modelling strategy is twofold.
+Firstly, a binary indicator variable will estimate the magnitude of effect the US presidential announcement and inauguration have on the US exchange rate.
+Secondly, the influence of a change in political party between the an US president and their predecessor will also be estimated.
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Magnitude of effect created by the helpline efforts can be estimated through linear models.
+Gaussian regression model is best suited for our analytical purposes.
+Other generalized linear models such as Logistic and Multilevel are not considered due to the nature of the variables in interest.
+Logistic regression model requires data that yields either an **yes** or **no** result, whereas multilevel model require more complexity between variables.
+Due to the nature of the Poisson function, the Poisson model requires count variable not available for our dataset.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
 ## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
 
@@
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
 

### 2024-04-15 scripts/04-model.R

@@ -31,28 +31,12 @@ inaug_model <-
   )
 
 
-# Attempt alternative RDD approach, ABANDONED
-#inaug_period_model <-
-#  stan_glm(
-#    formula = exchange_rate ~ date + inauguration_period,
-#    data = inaug_period_exchange,
-#    family = gaussian(),
-#    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-#    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-#    prior_aux = exponential(rate = 1, autoscale = TRUE),
-#   seed = 21
-#  )
+
+
 
 #### Save model ####
 saveRDS(
   inaug_model,
   file = "models/inaug_model.rds"
 )
-
-
-# alternative approach, ABANDONED
-#saveRDS(
-#  inaug_period_model,
-#  file = "models/inaug_period_model.rds"
-#)
 

### 2024-04-15 paper/paper.qmd
finished drafting the data section

@@
 This paper will focus on one particular aspect that contributes to the change in USD exchange rate: the expectation of the populous on the future of the United States, perceived through the results of the US Presidential Election.
+The election of a new president is often one of the biggest political shift for a country, shifts in stratergies and focus will cause many people to reevaluate their faith for their country.
 
@@
 
+The DEXCAUS [@Federal_Reserve] dataset includes the daily exchange rate between USD and CAD (base USD = 1) from 1971-01-04 to\
+2024-03-29 (date the dataset is downloaded for the purpose of the paper).
+Raw data contains only the date and the corresponding exchange rate of the day.
+Exchange rate in this dataset is used as the estimate of this paper.
+
 ```{r}
@@
 
-## Data on Election Results
+## Data on Election Results {#sec-election}
 
-### Selection of Data Source
+In order to measure the effect that the population's faith to their country have on exchange rate, this paper focuses on one particular political event that will quickly change the population's expectation for their country in a short amount of time, the presidential election.
+Analysis will be focused on the election result announcement and the inauguration period (please see specific definition of the term inauguration period in the subset below @sec-inauguration).
+Data of past election results and inauguration dates are obtained through the US Census Bureau [@Census] and cross-referenced with Frank LaRose Ohio Secretary of State [@Ohio] for accuracy.
+The final dataset use the inauguration dates and political party information to determine the effect of having a new president on the USD exchange rate.
 
@@
 
-knitr::kable(cleaned_inauguration, col.names = c("President", "Party", "Inauguration Date", "Change Party"), align = c("c", "c", "c", "c"))
+select_inauguration <- cleaned_inauguration |> select(-change_party)
+
+knitr::kable(select_inauguration, col.names = c("President", "Party", "Inauguration Date"), align = c("c", "c", "c"))
 ```
 
-### Construction of Inauguration Week
+### Construction of Inauguration Week {#sec-inauguration}
+
+Using the inauguration date information present in the dataset, we will construct an announcement period, allowing the news of the result to sink in and give time for the exchange rate to react to the announcement.
+This period needs to be relatively short to ensure that the change in exchange rate is causal by the election result announcement and **only** the announcement.
+
+The final construction will be referred to as the *inauguration week* for the remainder of this paper.
+The inauguration week is a +/-3 days week centered on the original inauguration date for each past US president.
 $$Day1 \hspace{5mm} Day2 \hspace{5mm}  Day3 \hspace{7mm}  \textbf{Inauguration Day} \hspace{7mm} Day5 \hspace{5mm} Day6 \hspace{5mm} Day7$$
+
+## Final Dataset
+The final dataset combines the date and exchange rate columns from the DEXCAUS [@Federal_Reserve] dataset and uses the inauguration week (constructed through the inauguration dates from the election results dataset, @sec-election).
+The column *Inauguration Period* has a value of 1 if the date is within the inauguration week and 0 otherwise.
+A new column *Change Party* is added, it receives a value of 1 if the political party of the new president is different from the previous president's and 0 otherwise.
 ```{r}
@@
 
-## Explaination for small R^2
+## Explaination for small R\^2
+
 ```{r}
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 

### 2024-04-16 paper/paper.qmd
re-rendered the paper.pdf for peer review

@@
 
-## Data on Exchange Rate
-
-The DEXCAUS [@Federal_Reserve] dataset includes the daily exchange rate between USD and CAD (base USD = 1) from 1971-01-04 to\
-2024-03-29 (date the dataset is downloaded for the purpose of the paper).
-Raw data contains only the date and the corresponding exchange rate of the day.
-Exchange rate in this dataset is used as the estimate of this paper.
-
 ```{r}
@@
   geom_line() + theme_minimal() +
-  labs(
-    x = "Year",
-    y = "Exchange Rate: USD to CAD (Base Currency: USD)",
+  labs(x = "Year", y = "Exchange Rate: USD to CAD (Base Currency: USD)",
     title = "Exchange Rate from 1971 to 2024"
@@
 ```
+## Data on Exchange Rate
+
+The DEXCAUS [@Federal_Reserve] dataset includes the daily exchange rate between USD and CAD (base USD = 1) from 1971-01-04 to 2024-03-29 (date the dataset is downloaded for the purpose of the paper).
+Raw data contains only the date and the corresponding exchange rate of the day.
+Exchange rate in this dataset is used as the estimate of this paper.
 
@@
 In order to measure the effect that the population's faith to their country have on exchange rate, this paper focuses on one particular political event that will quickly change the population's expectation for their country in a short amount of time, the presidential election.
-Analysis will be focused on the election result announcement and the inauguration period (please see specific definition of the term inauguration period in the subset below @sec-inauguration).
+Analysis will be focused on the election result announcement and the inauguration period (please see specific definition of the term inauguration period in the subsection below @sec-inauguration).
 Data of past election results and inauguration dates are obtained through the US Census Bureau [@Census] and cross-referenced with Frank LaRose Ohio Secretary of State [@Ohio] for accuracy.
@@
 ## Final Dataset
+
 The final dataset combines the date and exchange rate columns from the DEXCAUS [@Federal_Reserve] dataset and uses the inauguration week (constructed through the inauguration dates from the election results dataset, @sec-election).
@@
 A new column *Change Party* is added, it receives a value of 1 if the political party of the new president is different from the previous president's and 0 otherwise.
+
 ```{r}
@@
 
-## Third discussion point
-
 ## Weaknesses and next steps
@@
 # Appendix {.unnumbered}
-
-# Additional data details
 

### 2024-04-16 paper/paper.qmd
updated result section

@@
 
-Our results are summarized in @tbl-modelresults.
+@tbl-modelresults shows the coefficients of the predictor variables of our Model.
+We are concerned with the top half of the table, as it provides the values of the intercept representing the average exchange rate during non-inauguration weeks the as well as the coefficient representing effect size of inauguration periods and changes in political party.
+
+Looking at the values, we see that, outside of inauguration announcement weeks, the average exchange rate between USD and CAD is 1.23 between the years 1971 to 2024.
+When a new US presidental is anncouncement and subsequently inaugurated, an average of 0.14 drop occurs to the USD exchange rate.
+However, looking at the third coefficient, the 0.18 coefficient for the *change_party* variable suggests that, if a change of political party occurred during the current election season, then the US exchange rate will increase by 0.18 unit.
+This suggests that the net effect of announcing a new president who comes from a different political party than their immediate predecessor will, on average, increase the USD exchange rate by 0.04.
 
@@
 ## Explaination for small R\^2
+The model summary table @tbl-modelresults shows an extremely small R^2 value.
+The R^2 value is traditioally used to indicates the explanatory power of the model, that is, how trustworthy the model is on estimating the relationship between the estimand and the estimate.
+
+Many factors contributes to the R^2 value, one of the most prominent factor in our case is the effect size of our estimand.
+As mentioned in the introduction section, the estimand of this paper is how the announcement and inauguration of a new US president, and how a change in political party, influence the USD exchange rate.
+There are 13358 observations in the final dataset used for modelling, out of all of these observations, 40 of them contributes to the measurement of the effect of inauguration anncouncement, and 30 on the effect of a change in political party.
+In a practical sense, only 0.3% of the data are used to estimate the effect of the inauguration anncouncment and even less on the change of party.
+This contributes the the low R^2 value significantly.
+Dispite this, the model has still proven to be a good fit (see @sec-model-details), the predictors are still valuable despite the low R-squared.
+@fig-size_visual visualizes the effect size (in blue) and the remaining oberservations outside of the inauguration periods.
 
@@
 #| label: fig-size_visual
-#| fig-cap: Estimate average change in Exchange Rate
+#| fig-cap: Effect size visualization
 #| echo: false

### 2024-04-16 paper/paper.qmd
updated theme on fig-size_visual

@@
     formula = "y ~ x"
-  ) + scale_color_manual(values = colors) +
+  ) + scale_color_manual(values = colors) + theme_minimal() +
   labs(

### 2024-04-16 paper/paper.qmd
newst rendering of paper.pdf, with updated fig-size_visual

--- 2024-04-16 paper/paper.qmd
@@
   )) +
-  geom_point(alpha = 0.005) +
+  geom_line(alpha = 0.1) +
   geom_smooth(

### 2024-04-17 paper/paper.qmd
deleted an unused folder and update READMe and title of paper

@@
 ---
-title: "Effect of Election on Currency Fluctuation"
+title: "USD Exchange Rate decreases during Election Anncouncement Period"
 subtitle: "Analysis of USD Exchange Rate Fluctuation during Inaugurtion week"
@@
 
-Weaknesses and next steps should also be included.
+- small effect size causing low explanitory power
+- inauguration dates does NOT include presidential announcement results where the previous president gets re-elected for the next season, will cause bias in estimation.
 

### 2024-04-17 paper/paper.qmd
corrected multiple spelling mistakes in the paper

--- 2024-04-16 paper/paper.qmd
@@
 ---
-title: "Effect of Election on Currency Fluctuation"
-subtitle: "Analysis of USD Exchange Rate Fluctuation during Inaugurtion week"
+title: "USD Exchange Rate decreases during Election Anncouncement Period"
+subtitle: "Analysis of USD Exchange Rate Fluctuation during Inauguration week"
 author: 
@@
 We find that election result announcement has an average negative effect on USD exchange rate whereas a change of political party contributes positively with a slightly larger magnitude.
-Our result suggests that exchange USD for another currency closly after a Presidential election with a change of political party could be profitable."
+Our result suggests that exchange USD for another currency closly after a Presidential election with a change of political party could be profitable, therefore it might be profitable to buy in USD during election result anncouncement weeks depending on the political party of the new president."
 format: pdf
@@
 
-USD is one of the most popular currency in circulation today.
+USD is one of the most popular currencies in circulation today.
 As a foreign buyer of USD product, the final price can vary greatly depending on the exchange rate between one's local currency and USD.
@@
 This paper will focus on one particular aspect that contributes to the change in USD exchange rate: the expectation of the populous on the future of the United States, perceived through the results of the US Presidential Election.
-The election of a new president is often one of the biggest political shift for a country, shifts in stratergies and focus will cause many people to reevaluate their faith for their country.
+The election of a new president is often one of the biggest political shifts for a country, shifts in stratergies and focus will cause many people to reevaluate their faith for their country.
 
@@
 USD v.s.
-CAD exchange rate from 1971 to 2024 is pulled from the @FRED website, and past presidential election results and inauguration dates are gathered and double-checked through @Cencus and @Ohio.
+CAD exchange rate from 1971 to 2024 is pulled from the @FRED website, and past presidential election results and inauguration dates are gathered and double-checked through @Census and @Ohio.
 The estimand of the paper is how the announcement and inauguration of a new US president, and how a change in political party, influence the USD exchange rate.
@@
   geom_line() + theme_minimal() +
-  labs(x = "Year", y = "Exchange Rate: USD to CAD (Base Currency: USD)",
-    title = "Exchange Rate from 1971 to 2024"
+  labs(x = "Year", y = "Exchange Rate: USD to CAD (Base Currency: USD)"
   )
 ```
+
 ## Data on Exchange Rate
@@
 Looking at the values, we see that, outside of inauguration announcement weeks, the average exchange rate between USD and CAD is 1.23 between the years 1971 to 2024.
-When a new US presidental is anncouncement and subsequently inaugurated, an average of 0.14 drop occurs to the USD exchange rate.
+When a new US presidental is announcement and subsequently inaugurated, an average of 0.14 drop occurs to the USD exchange rate.
 However, looking at the third coefficient, the 0.18 coefficient for the *change_party* variable suggests that, if a change of political party occurred during the current election season, then the US exchange rate will increase by 0.18 unit.
@@
 
-## Explaination for small R\^2
-The model summary table @tbl-modelresults shows an extremely small R^2 value.
-The R^2 value is traditioally used to indicates the explanatory power of the model, that is, how trustworthy the model is on estimating the relationship between the estimand and the estimate.
+## Explanation for small R\^2
 
-Many factors contributes to the R^2 value, one of the most prominent factor in our case is the effect size of our estimand.
+The model summary table @tbl-modelresults shows an extremely small R\^2 value.
+The R\^2 value is traditionally used to indicates the explanatory power of the model, that is, how trustworthy the model is on estimating the relationship between the estimand and the estimate.
+
+Many factors contributes to the R\^2 value, one of the most prominent factor in our case is the effect size of our estimand.
 As mentioned in the introduction section, the estimand of this paper is how the announcement and inauguration of a new US president, and how a change in political party, influence the USD exchange rate.
-There are 13358 observations in the final dataset used for modelling, out of all of these observations, 40 of them contributes to the measurement of the effect of inauguration anncouncement, and 30 on the effect of a change in political party.
-In a practical sense, only 0.3% of the data are used to estimate the effect of the inauguration anncouncment and even less on the change of party.
-This contributes the the low R^2 value significantly.
-Dispite this, the model has still proven to be a good fit (see @sec-model-details), the predictors are still valuable despite the low R-squared.
-@fig-size_visual visualizes the effect size (in blue) and the remaining oberservations outside of the inauguration periods.
+There are 13358 observations in the final dataset used for modelling, out of all of these observations, 40 of them contributes to the measurement of the effect of inauguration announcement, and 30 on the effect of a change in political party.
+In a practical sense, only 0.3% of the data are used to estimate the effect of the inauguration announcement and even less on the change of party.
+This contributes the the low R\^2 value significantly.
+Despite this, the model has still proven to be a good fit (see @sec-model-details), the predictors are still valuable despite the low R-squared.
+@fig-size_visual visualizes the effect size (in blue) and the remaining observations outside of the inauguration periods.
 
@@
 
-Weaknesses and next steps should also be included.
+-   small effect size causing low explanatory power
+-   inauguration dates does NOT include presidential announcement results where the previous president gets re-elected for the next season, will cause bias in estimation.
 

### 2024-04-19 paper/paper.qmd
Completed the Discussion section and editted the READMe

@@
 The past US president inauguration dates are pulled from the US Census Bureau [@Census] and cross-referenced with Frank LaRose Ohio Secretary of State [@Ohio].
-
-Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], rstanarm [@rstanarm], arrow [@arrow], ggplot2 [@ggplot2], and knitr [@knitr], with additional help from Microsoft Excel [@citeExcel].
-Detailed description of each dataset can be found in the subsections below.
-
 ```{r}
@@
 ```
+Data is cleaned and analyzed using the open source statistical programming language R [@citeR] and supporting packages tidyverse [@tidyverse], janitor [@janitor], rstanarm [@rstanarm], arrow [@arrow], ggplot2 [@ggplot2], and knitr [@knitr], with additional help from Microsoft Excel [@citeExcel].
+Detailed description of each dataset can be found in the subsections below.
 
@@
 Exchange rate in this dataset is used as the estimate of this paper.
+See @fig-exchange for visualization of this dataset.
 
@@
 The final dataset use the inauguration dates and political party information to determine the effect of having a new president on the USD exchange rate.
+See @tbl-inauguration for a sample of the election result dataset.
 
@@
 
-The final dataset combines the date and exchange rate columns from the DEXCAUS [@Federal_Reserve] dataset and uses the inauguration week (constructed through the inauguration dates from the election results dataset, @sec-election).
+The final dataset, @tbl-exchange_inaug, combines the date and exchange rate columns from the DEXCAUS [@Federal_Reserve] dataset and uses the inauguration week (constructed through the inauguration dates from the election results dataset, @sec-election).
 The column *Inauguration Period* has a value of 1 if the date is within the inauguration week and 0 otherwise.
@@
 Looking at the values, we see that, outside of inauguration announcement weeks, the average exchange rate between USD and CAD is 1.23 between the years 1971 to 2024.
-When a new US presidental is announcement and subsequently inaugurated, an average of 0.14 drop occurs to the USD exchange rate.
+When a new US president is announced and subsequently inaugurated, an average of 0.14 drop occurs to the USD exchange rate.
+In practical terms, this means that a 11% drop from the average exchange rate of 1.23 occured in the span of 1 week due to the announcement.
+
 However, looking at the third coefficient, the 0.18 coefficient for the *change_party* variable suggests that, if a change of political party occurred during the current election season, then the US exchange rate will increase by 0.18 unit.
@@
 ## Expectation of the population
+Population expectation plays a significant role as a key driver of exchange rate fluctuations in the field of economics.
+However, the term expectation is often ambiguous and hard to quantify.
+This paper seeks to explore the tangible impact of expectations on exchange rates by focusing on a specific, highly influential event: the U.S.
+presidential election.
+This particular event was chosen for its prominence and its profound effect on national sentiment.
+It stands out due to its high engagement levels among U.S.
+citizens, with a notable participation rate of 66.8% among eligible voters aged 18 and older in the 2020 presidential election, according to data referenced from @2020Election (surveys conducted by @Census2020Data).
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The presidential election is an optimal choice for analysis not only because of its widespread recognition and significant voter turnout but also due to its measurable impact on economic indicators such as the exchange rate.
+As detailed in the section @sec-results of this paper, there is an observable pattern related to changes in political leadership and their effects on currency valuation.
+Specifically, the inauguration week of a new president who belongs to the same political party as their immediate predecessor typically sees an decrease in the exchange rate by 11% compared to the exchange rate at non-inauguration weeks.
+Conversely, a change in the ruling party correlates with an average net increase of 3.25% in the exchange rate.
+Viewing an increase in the exchange rate as an indicator of growing confidence in the country's direction, these findings suggest several implications about public sentiment and economic expectations.
 
-## Change of Political Party
+## Ideology of Political Parties
+
+During election seasons, the focus isn't just on the incumbent president but also heavily on the speeches and promises made by presidential candidates.
+These candidates not only present their personal ideologies but also propose new policies and legislative changes they plan to implement if elected.
+
+Candidates from the same political party often share core values and ideologies, yet they might differ in their specific priorities and proposed policies.
+For instance, it's common for all Democratic candidates to support initiatives related to refugee and immigration issues, although their specific approaches to these issues can vary.
+When candidates from the same party are elected, it typically indicates that they will continue the country's current trajectory, particularly in how the U.S.
+handles international affairs.
+Electing a candidate from the same party as the incumbent suggests that voters are generally pleased with the current direction of the nation and wish to maintain continuity.
+
+On the other hand, when a candidate from a different political party is elected, it usually signifies that a significant portion of the electorate believes the country needs a new direction.
+This desire for change is reflected in the public's support for the candidate's proposed shifts in political focus and new policies.
+This can lead to a change in how the U.S.
+is perceived both domestically and internationally.
+
+The election of a president from a different party than their predecessor often results in an increase in the USD exchange rate.
+This phenomenon can occur because the financial markets react to the potential for substantial policy changes that may foster economic growth or improve the fiscal outlook of the country.
+Investors might anticipate that new policies will be beneficial for business and economic stability, which can lead to increased foreign investment and a stronger dollar.
+The promise of a fresh approach and new leadership can instill confidence among investors and the business community, leading to positive fluctuations in the currency's value.
+
+## External Shifts
+The influence of expectation extends beyond just the domestic population of the United States.
+It also significantly affects foreign stakeholders such as holders of U.S.
+government bonds and owners of international companies that contribute to the U.S.
+GDP.
+These groups are deeply interested in the stability and policies of the U.S.
+economy because changes can directly impact their investments and business operations.
+As such, their perceptions and expectations about the U.S.
+economy can lead to fluctuations in the USD exchange rates.
+
+Furthermore, the inauguration of a new president can lead to shifts in international relations and economic policies.
+This change in leadership can create uncertainty about future policies, which might affect existing collaborations and agreements made during the previous administration.
+For example, a new president might reconsider or revise international trade agreements, which could potentially disrupt established economic partnerships.
+Such uncertainties may lead to instability in exchange rates as investors and foreign businesses try to predict and adapt to the new policy environment.
+Political dynamics in the United States can have a far-reaching impact on global economic interactions and the overall behavior of the USD on the international stage.
 
 ## Weaknesses and next steps
+The proposed model, while insightful, has several limitations that need addressing to enhance its accuracy and applicability.
+One of the primary weaknesses is its small effect size, which contributes to its low explanatory power, making it less effective in predicting USD exchange rate fluctuations around U.S.
+presidential inaugurations.
+Additionally, the model does not account for inauguration dates following the re-election of a sitting president, which can introduce bias into the estimations.
+Furthermore, it overlooks the constitutional limit that prevents a president from being elected more than twice consecutively, which could skew results where such scenarios are relevant.
 
--   small effect size causing low explanatory power
--   inauguration dates does NOT include presidential announcement results where the previous president gets re-elected for the next season, will cause bias in estimation.
+Moving forward, the next steps to refine the model include a comprehensive collection of all election announcement dates for past U.S.
+presidents who have been re-elected consecutively.
+This data will be integrated into the model to provide a more accurate reflection of political impacts on economic indicators.
+Additionally, it's crucial to isolate and eliminate the influence of other significant events that historically affect the USD exchange rate.
+By addressing these points, the data can offer more reliable and significant insights into the dynamics between presidential elections and economic fluctuations.
 
